---
title: "Assignment 4"
author: "Cristina Davila Gonzalez"
date: "10/4/2020"
output: html_document
---

```{r, message=FALSE}
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
library(dplyr)
library(plyr)
```

## Load Locations

In this exercise, I will look at hubway stations and landmarks in Downtown Boston, MA. 

```{r, message=FALSE}
hubway_stations <- st_read(
  "http://bostonopendata-boston.opendata.arcgis.com/datasets/ee7474e2a0aa45cbbdfe0b747a5eb032_0.kml?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D")

landmarks <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/7a7aca614ad740e99b060e0ee787a228_3.kml")

```

## Street data: Downtown Boston, MA.

Street network data from Open Street Map
```{r, message=FALSE}
opq(bbox = 'Downtown Boston MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_xml(file = 'OTP/graphs/default/downtownb_streets.osm')
```


SF features from Open Street Map to plot maps.
```{r, message=FALSE}
DB_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

downtown_street_features <- opq(bbox = 'Downtown Boston MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_sf()

downtown_streets <- downtown_street_features$osm_lines %>%
  st_transform(crs = DB_state_plane)
```

Plotting downloaded streets
```{r fig.height=2, fig.width=2}
ggplot(downtown_streets) +
  geom_sf() +
  theme_map()+
   annotation_north_arrow(location = "tl",   height = unit(0.7, "cm"),
  width = unit(0.7, "cm") )
```

OTP Setup
```{r message=FALSE, results='hide'}
path_otp <- otp_dl_jar("OTP")
```

Street and transit networks
```{r, message=FALSE, results='hide'}
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")

otp_build_graph(otp = path_otp, dir = path_data, memory = 1024) 
```

```{r, message=FALSE, results='hide'}
otp_setup(otp = path_otp, dir = path_data, memory =1024)
```

Connecting to OpenTripPlanner
```{r, message=FALSE}
otpcon <- otp_connect()
```


## Isochrones and graphs: Areas within a five-minute **walk** and a five-minute **bike ride**

### Isochrone 1: default map type

```{r, message=FALSE}
iso_5min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = hubway_stations, 
                mode = "WALK", cutoffSec = 300) %>%
  st_transform(crs = DB_state_plane) %>%
  mutate(mode = "walk")

iso_5min_bike <- 
  otp_isochrone(otpcon = otpcon, fromPlace = hubway_stations, 
                mode = "BICYCLE", cutoffSec = 300) %>%
  st_transform(crs = DB_state_plane) %>%
  mutate(mode = "bike")

iso_all_modes <- rbind(iso_5min_bike, iso_5min_walk)

otp_stop()
```

```{r fig.height=5, fig.width=5}
right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 1, progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.2) +
  geom_sf(data = hubway_stations) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 5 minutes",
                       labels = c("By bike", "By foot")) +
  annotation_scale(location = "br")+
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```

### Isochrone 2: cartolight

```{r fig.height=5, fig.width=5}
ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 1, type= "cartolight", progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.2) +
  geom_sf(data = hubway_stations) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 5 minutes",
                       labels = c("By bike", "By foot"),
                       option = "plasma") +
  annotation_scale(location = "br")+
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```

### Scatterplot: Comparisons between walkshed and bikeshed isochrone areas 

```{r, message=FALSE}
iso_areas <- iso_all_modes %>%
  mutate(area = st_area(iso_all_modes)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area) 

ggplot(iso_areas, 
       aes(x = as.numeric(walk), y = as.numeric(bike))) +
  geom_point() +
  stat_smooth(color = "black", linetype = 2, size = 0.5) +
  scale_x_continuous(name = 
            "Area within a five-minute walking distance of a hubway station\n(square km)",
            breaks = breaks <- seq(10000, 130000, by = 20000),
            labels = breaks / 1000000) +
  scale_y_continuous(name = 
            "Area within a five-minute biking distance of a hubway station\n(square km)",
            breaks = breaks <- seq(0, 700000, by = 100000),
            labels = breaks / 1000000) +
  ggtitle("Downtown Boston", subtitle = paste0("hubway stations"))+
  theme_pander()
```

## Creating Map of landmarks within 5 min bike ride of a hubway station in Downtown Boston

### Selecting landmarks within 5 min bikeshed of hubway stations
```{r}
landmarks <- landmarks %>% 
  st_transform(DB_state_plane)

landmarks <- landmarks[iso_5min_bike,]
ggplot(iso_5min_bike) +
  geom_sf(color= NA) +
  geom_sf(data = landmarks, 
          fill = "black", color= NA) +
  theme_map()
```

### Mapping landmarks within 5 min bike ride of a hubway station in Downtown Boston
```{r, fig.height=8, fig.width=8}
right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_5min_bike) +
  annotation_map_tile(zoomin = 1, type = "cartolight", progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.2) +
  geom_sf(data = landmarks, fill= "black", color= NA, size = 1.5) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 5 minutes",
                       labels = c("By bike")) +
  annotation_scale(location = "br")+
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```

```{r}
otp_stop()
``` 

