---
title: "Assignment 1"
author: "Cristina Davila Gonzalez"
date: "9/13/2020"
output: html_document
---
# VIS Assignment 1

For the purposes of focusing this exercise in **experimenting more with ggplot**, I'll be using Taelor Malcolm's variables in her **Week 2** assignment. In my particular case, I'll be applying them in **Honolulu County, Hawaii**. First, I'm loading all the libraries I'll be needing in this case:

```{r load libraries, message = FALSE}
library(tidycensus)
library(tidyverse)
library (ggplot2)
```

```{r}
vars <- c("B06009_004",
          "B25001_001",
          "B25003_003",
          "B06007_002",
          "B01002_001",
          "B01003_001",
          "B03002_003",
          "B03002_004",
          "B03002_006",
          "B03002_012",
          "B01001_026",
          "B06010_003",
          "B06010_004",
          "B06010_005",
          "B06010_006",
          "B06010_007",
          "B06010_008",
          "B06010_009",
          "B06010_010",
          "B06010_011")
```

## Loading 2014-2018 5 year ACS data for Honolulu County, HI

```{r,  message=FALSE, results='hide'}
Honolulu_tracts <- get_acs(geography = "tract", 
                       variables = vars,
                       state = "HI", 
                       county = "Honolulu",
                       surver= "acs5",
                       output = "wide") %>% 
  
             rename(tot_bach = "B06009_004E",
                    tot_hh = "B25001_001E",
                    tot_rent_occ = "B25003_003E",
                    tot_eng = "B06007_002E",
                    med_age = "B01002_001E",
                    tot_pop = "B01003_001E",
                    pop_inc = "B06010_003E",
                    bel_10 = "B06010_004E", 
                    bel_15 = "B06010_005E",
                    bel_25 = "B06010_006E",
                    bel_35 = "B06010_007E",
                    bel_50 = "B06010_008E",
                    bel_65 = "B06010_009E",
                    bel_75 = "B06010_010E",
                    abo_75 = "B06010_011E",
                    nh_white_pop = "B03002_003E",
                    nh_black_pop = "B03002_004E",
                    nh_asian_pop = "B03002_006E",
                    hispanic_pop = "B03002_012E") %>%
              mutate(pct_rent_occ = tot_rent_occ / tot_hh,
                     pct_bach = tot_bach / tot_pop,
                     pct_eng = tot_eng / tot_pop,
                     majority_inc = case_when(
                     (bel_10 + bel_15 + bel_25) / pop_inc > 0.5 ~ "bel_25",
                     (bel_35 + bel_50 + bel_65 + abo_75) / pop_inc > 0.5 ~ "abo_25",
                     (bel_10 + bel_15 + bel_25 + bel_35 + bel_50 + bel_65 +
                      bel_75 + abo_75)/pop_inc < 0.5 ~ "other", 
                     TRUE ~ "None"),
                     majority_race_eth = case_when(
                     nh_white_pop / tot_pop > 0.5 ~ "nh_white",
                     nh_black_pop / tot_pop > 0.5 ~ "nh_black",
                     nh_asian_pop / tot_pop > 0.5 ~ "nh_asian",
                    hispanic_pop / tot_pop > 0.5 ~ "hispanic",
                    (nh_white_pop + nh_black_pop + nh_asian_pop + hispanic_pop) /
                     tot_pop < 0.5 ~ "other",
                     TRUE ~ "None")) %>%
              filter(tot_pop > 0, tot_hh > 0) %>%
              select(GEOID, pct_rent_occ, pct_bach, pct_eng,
                     majority_inc, med_age, majority_race_eth, tot_pop)
```

I decided to incorporate the Wes Anderson color palette, which I got from a shared link shared in a R Zoom study session with fellow incoming MUPs.

```{r}
library(wesanderson)

```


## Plot 1: Pie chart

## Majority race/ethnicity

```{r, message=FALSE, results='hide'}
pie_chart <- ggplot(Honolulu_tracts, aes(x = 1, fill = majority_race_eth)) +
  geom_bar(position = "fill") +
  coord_polar(theta = "y") +
  scale_y_continuous(name = "Number of tracts") +
  scale_fill_manual (name = "Majority race/ethnicity",
                     values = wes_palette("Rushmore1", n = 4),
                       labels = c("Hispanic",
                                  "Non-Hispanic\nAsian",
                                  "Non-Hispanic\nwhite",
                                  "No majority")) +
  theme_void() 

pie_chart
```

## Plot 2: Area chart

## Median age within tract

```{r, message=FALSE, results='hide'}

ggplot(Honolulu_tracts, aes(x=med_age)) +
  geom_density(fill="#CAAC3A", color="#e9ecef", alpha=0.8) +
  scale_x_continuous(name = "Median age within tract",
                     breaks = seq(10, 80, by = 10)) +
  
  theme_minimal()

```

## Plot 3: Scatterplot with overlapping points chart

## Median age within tract and race/ethnicity

```{r, message=FALSE, results='hide'}

ggplot(Honolulu_tracts, 
       aes(x = med_age,
           y = majority_race_eth,
           color= majority_race_eth, )) +
  geom_point(position = "jitter",
             size = 1,
             alpha = 3) +
  stat_smooth(color = "#69b3a2", linetype = 1, size = 1) +
  scale_x_continuous(name = "Median age within tract") +
  scale_y_discrete(name = element_blank(),
                       labels = c("Hispanic",
                                  "Non-Hispanic\nAsian",
                                  "Non-Hispanic\nwhite",
                                  "No majority")) +
   scale_color_manual (name = "Majority race/ethnicity",
                     values = wes_palette("GrandBudapest1", n = 4),
                       labels = c("Hispanic",
                                  "Non-Hispanic\nAsian",
                                  "Non-Hispanic\nwhite",
                                  "No majority")) +
  
  theme_minimal()+ 
  theme(legend.position ="none")
```

I got the #69b3a2 color for the lines between dots from Taelor's file. 

## Plot 4: Violin and scatter chart

## Percent of individuals that only speak English, and majority race/ethnicity

```{r, message=FALSE, results='hide'}

ggplot(Honolulu_tracts, aes(x = majority_race_eth, y = pct_eng, color = majority_race_eth)) +
  geom_violin() +
  geom_point(position = "jitter",
             alpha = 0.5,
             size = 0.5) +
  scale_x_discrete(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic white",
                                  "No majority")) +
  scale_y_continuous(name = "Percent of individuals that only speak English",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_color_manual(name = "Majority race/ethnicity",
                     values = wes_palette("GrandBudapest2", n = 4),
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic White",
                                  "Non-Hispanic Black", 
                                  "No Majority")) +
  theme_minimal() +
  theme(legend.position = "none")
  
```

## Plot 5: Bar chart

## Race/ethnicity and income

```{r, message=FALSE, results='hide'}
ggplot(Honolulu_tracts, aes(x = majority_race_eth,fill = majority_inc)) +
  geom_bar() +
  scale_y_continuous(name = "Number of tracts") +
  scale_x_discrete(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic\nAsian",
                                  "Non-Hispanic\nwhite",
                                  "No majority")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 3),
                    name = element_blank(),
                        labels = c("Income Below $25,000", "Income Above $25,000")) +

  theme_minimal() +
  theme(legend.key.size = unit(0.1, "cm"))
```

## Plot 6: Counts chart

## Percent of renter occupied units, race/ethnicity and majority income

```{r, message=FALSE, results='hide'}

ggplot (Honolulu_tracts) +
  geom_point(mapping = aes(x = majority_race_eth, y = pct_rent_occ, size= majority_inc, color= majority_inc))+
  
 scale_size_discrete(name = "Majority income",
                     labels = c("NA", "Income Below $25,000", "Income Above $25,000")) +
 scale_y_continuous(name = "Percent of renter occupied units",
                     breaks = seq(0, 1, by = 0.25),
                     labels = paste(seq(0, 100, by = 25), "%", sep = "")) +
 scale_x_discrete(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic White",
                                  "Non-Hispanic Black", 
                                  "No Majority")) +
  scale_color_manual(name = "Majority income",
                     values = wes_palette("FantasticFox1", n = 3),
                     labels = c("NA", "Income Below $25,000", "Income Above $25,000")) +
  theme_minimal() +
  theme(legend.key.size = unit(0.1, "cm"))

```
## Plot 7: 2D Density chart

## Percent of renter occupied units, Median age within tract and race/ethnicity

developed with the additional support from Julia Meinhardt, MUP'22.

```{r, message=FALSE, results='hide'}

ggplot(Honolulu_tracts) +
    geom_density2d(mapping= aes(y = pct_rent_occ,
           x = med_age,
           color = majority_race_eth)) +
  scale_y_continuous(name = "Percent of renter occupied units",
                     breaks = seq(0, 1, by = 0.25),
                     labels = paste(seq(0, 100, by = 25), "%", sep = "")) +
  scale_x_continuous(name = "Median age within tract",
                     breaks = seq(10, 80, by = 10)) +
  scale_color_manual(name = "Majority race/ethnicity",
                     values = wes_palette("Darjeeling2", n = 4),
                     labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic White",
                                  "Non-Hispanic Black", 
                                  "No Majority")) +
  theme_minimal() 
  theme(legend.key.size = unit(0.1, "cm"))
```
## Plot 8: Shapes scatter plot chart

## Percent of individuals that only speak English, income and median age within tract

```{r, message=FALSE, results='hide'}

ggplot (Honolulu_tracts) +
        geom_point(mapping = aes(y = pct_eng, x = med_age, shape = majority_inc, color=majority_inc )) +
     scale_y_continuous(name = "Percent of individuals that only speak English",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
   scale_x_continuous(name = "Median age within tract",
                     breaks = seq(10, 80, by = 10)) +
   scale_shape_discrete(name = "Majority income",
                     labels = c("NA", "Income Below $25,000", 
                                "Income Above $25,000")) + 
   scale_color_manual(name = "Majority income",
                     values = wes_palette("Moonrise3", n = 3),
                     labels = c("NA", "Income Below $25,000", "Income Above $25,000")) +
    theme_minimal() 
    theme(legend.key.size = unit(0.1, "cm"))
```

## Plot 9: Fit Line scatter plot Chart 

## Total population and percent of people with a bachelor's degree

```{r, message=FALSE, results='hide'}

ggplot(Honolulu_tracts, aes(y=pct_bach, x=tot_pop)) +
  geom_point(mapping = aes(color = majority_race_eth)) + 
  geom_smooth(fill="gray", color="#69b3a2", alpha=0.3, size = 0.5) +
  scale_y_continuous(name = "Percent with bachelor's degree",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = ""))  +
  scale_x_continuous(name = "Total Population")+
  scale_color_manual(name = "Majority race/ethnicity",
                     values = wes_palette("Darjeeling1", n = 4),
                     labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic White",
                                  "Non-Hispanic Black", 
                                  "No Majority"))+
  theme_minimal()
  theme(legend.key.size = unit(0.1, "cm"))
```


## Plot 10: Fit Line scatter plot Chart within a polar coordinate system

## Percent of individuals that only speak English, percent of renter occupied units and majority race/ethnicity

```{r, message=FALSE, results='hide'}
ggplot(Honolulu_tracts, 
       aes(x = pct_rent_occ,
           y = pct_eng,
           color= majority_race_eth )) +
  geom_point(alpha = 1, size = 0.5) +
  stat_smooth(color = "#69b3a2", linetype = 5, size = 0.5) +
  scale_x_continuous(name = "Percent of renter occupied units",
                     breaks = seq(0, 1, by = 0.1),
                    labels = paste(seq(0, 100, by = 10), "%", sep = ""))+
   scale_y_continuous(name = "Percent of individuals that only speak English",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_color_manual(name = "Majority race/ethnicity",
                     values = wes_palette("BottleRocket2", n = 4),
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic White",
                                  "Non-Hispanic Black", 
                                  "No Majority"))+
  coord_polar(theta = "x") +
  theme_minimal() 
  theme(legend.key.size = unit(0.1, "cm"))

```
