---
title: "Assignment 2"
author: "Cristina Davila Gonzalez"
date: "9/20/2020"
output: html_document
---
## Libraries

```{r, message=FALSE , results='hide'}
library(ggplot2)
library(sf)
library(tidyverse)
library(ggspatial)
library(cartogram)
library(ggthemes)
library(wesanderson)
library(viridis)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)

```

# World map

```{r, results='hide'}
world <- ne_countries(scale = 50, returnclass = "sf") 
world_map <- ggplot(world) +
  geom_sf(fill = NA, color = "#5498ae") +
  theme_minimal()

world_map

```

# South America
```{r, results='hide'}
sa_countries <- world %>%
  filter(continent == "South America")

sa_map <- ggplot(sa_countries) +
  geom_sf(fill = NA, color = "#85b6c3") + 
  theme_minimal()

sa_map
```

## South America in 3 map projections

```{r, results='hide'}
lcc69_sa <- "+proj=lcc +lat_1=-5 +lat_2=-42 +lat_0=-32 +lon_0=-60 +x_0=0 +y_0=0 +ellps=aust_SA +units=m +no_defs"
laea_sa <- "+proj=laea +lat_0=-10 +lon_0=-70 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"
br_poly <- "+proj=poly +lat_0=0 +lon_0=-54 +x_0=5000000 +y_0=10000000 +ellps=aust_SA +units=m +no_defs"
```

```{r, results='hide'}
ggplot(sa_countries) +
  geom_sf( fill = NA, color = "#ebcf58") +
  ggtitle("South America", subtitle = paste0("Map Projection 1: South America 1969 Lambert Conformal Conic")) +
  coord_sf(crs = lcc69_sa) +
  theme_minimal() 
```
```{r, results='hide'}
ggplot(sa_countries) +
  geom_sf(fill = NA, color = "#dd4d3c") +
  ggtitle("South America", subtitle = paste0("Map Projection 2: Lambert Azimuthal Equal Area (SERENA)")) +
  coord_sf(crs = laea_sa) +
  theme_minimal()
```
```{r, results='hide'}
ggplot(sa_countries) +
  geom_sf( fill = NA, color = "#7895d0") +
  ggtitle("South America", subtitle = paste0("Map Projection 3: Brazil Polyconic")) +
  coord_sf(crs = br_poly) +
  theme_minimal()
```

## South America map with transformed coordinate system: Brazil Polyconic
```{r, results='hide'}
sa_transformed <- sa_countries %>%
  st_transform(crs = br_poly)
```

## Representing Data in South America

For this assignment, I wanted to practice with both existing and imported datasets, so I generated graphics from the **rnaturalearth** library and **province and country-level data on COVID-19 from John Hopkins University’s Center for Systems Science and Engineering**, experimenting with both Wes Anderson and viridis palettes, respectively.

### GDP in South America: Chloropleth map

```{r, results='hide'}
options(scipen = 999)
pal <- wes_palette("Zissou1", 100, type = "continuous")
ggplot(data = sa_transformed) +
  ggtitle("South America") +
 geom_sf(color= NA, aes(fill = gdp_md_est)) +
     scale_fill_gradientn(colours = pal,
                         name = "Estimated Country GDP\n")+
  theme_map() +
  theme(legend.background = element_blank())
```

### Non-continuous cartogram

```{r, results='hide'}
pal <- wes_palette("Zissou1", 100, type = "continuous")

sa_cartogram_ncont <- sa_transformed %>%
  cartogram_ncont("gdp_md_est")

ggplot(sa_cartogram_ncont) +
  ggtitle("South America") +
  geom_sf(color = NA, aes (fill= gdp_md_est)) +
    scale_fill_gradientn(colours = pal,
                         name = "Estimated Country GDP\n")+
  theme_map() +
  theme(legend.background = element_blank())
```

### Dorling cartogram

```{r, results='hide'}
pal <- wes_palette("Zissou1", 100, type = "continuous")

sa_cartogram_dorling <- sa_transformed %>%
  cartogram_dorling("gdp_md_est")

ggplot(sa_cartogram_dorling) +
  ggtitle("South America") +
  geom_sf(color = NA, aes (fill= gdp_md_est)) +
     scale_fill_gradientn(colours = pal,
                         name = "Estimated Country GDP\n")+
  theme_map() +
  theme(legend.background = element_blank(),legend.position = "left"  )
```

## Mapping COVID-19 in South America - Importing and transforming data 

I found out how to upload data directly from **John Hopkins University COVID-19 Github repository** without the need to have the file in my computer, so I decided to give it a try, using data from most recent date I had when I began this exercise. 

I encountered myself with province-level data coexisting with country-level data, and Ignacio Lafuente kindly shared with me a way to not only group it so my graphs show the **total amount of cases per country**, but also to represent date on a scale of **numbers per million inhabitants**, to optimize interpretation:

```{r, results='hide'}
covid_world <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-19-2020.csv")

covid_world_2 <- select(covid_world, Country_Region, Confirmed)

covid_world_3 <- aggregate(covid_world_2[-1], by = list(covid_world$Country_Region), FUN = sum)

covid_sa <- sa_transformed %>%
  left_join(covid_world_3, by = c("name" = "Group.1")) %>%
  select(name, pop_est, Confirmed)

covid_sa$pop_est2 <-as.numeric(covid_sa$pop_est)

covid_sa$confirmed_per_m <- covid_sa$Confirmed / covid_sa$pop_est2 *1000000  

```

### Confirmed Cases in South America - Chloropleth map

```{r, results='hide'}
ggplot(covid_sa, aes(fill = confirmed_per_m)) +
   ggtitle("South America", subtitle = paste0("September 19th, 2020")) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "Number of reported\nCOVID-19 confirmed cases\nper million inhabitants",
    breaks = seq(0, 25000, by = 5000),
    labels = formatC(seq(0, 25000, by = 5000), 
                     format = "f", digits = 0)) +
  theme_map() +
  theme(legend.background = element_blank(), legend.position = "right")
```

### Confirmed Cases in South America - Dorling Cartogram

```{r, results='hide'}
covid_cartogram_dorling <- covid_sa %>%
  cartogram_dorling("pop_est2")

ggplot(covid_cartogram_dorling, aes(fill = confirmed_per_m)) +
  ggtitle("South America", subtitle = paste0("September 19th, 2020")) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "Number of reported\nCOVID-19 confirmed cases\nper million inhabitants",
    breaks = seq(0, 25000, by = 5000),
    labels = formatC(seq(0, 25000, by = 5000), 
                     format = "f", digits = 0)) +
  theme_map() +
  theme(legend.background = element_blank(), legend.position = "right")

```

### Confirmed Cases in South America - Non-continuous Cartogram

```{r, results='hide'}
covid_cartogram_ncont <- covid_sa %>%
  cartogram_ncont("pop_est2")

ggplot(covid_cartogram_ncont, aes(fill = confirmed_per_m)) +
  geom_sf(color = NA) +
   ggtitle("South America", subtitle = paste0("September 19th, 2020")) +
  scale_fill_viridis_c(
    name = "Number of reported\nCOVID-19 confirmed cases\nper million inhabitants",
    breaks = seq(0, 25000, by = 5000),
    labels = formatC(seq(0, 25000, by = 5000), 
                     format = "f", digits = 0)) +
 
  theme_map() +
  theme(legend.background = element_blank(), legend.position = "right")

```

### Confirmed Cases in South America - Proportional Symbol Map

```{r, results='hide'}
covid_centeroids <- covid_sa %>%
  st_centroid()

ggplot(sa_transformed) +
  geom_sf(fill = NA, color = "#67c562") +
   ggtitle("South America", subtitle = paste0("September 19th, 2020")) +
  geom_sf(data = covid_centeroids, 
          aes(size = confirmed_per_m),
          alpha = 0.5, color = "#433d80") +
  scale_size_continuous( name = "Number of reported\nCOVID-19 confirmed cases\nper million inhabitants",
    breaks = seq(0, 25000, by = 5000),
    labels = formatC(seq(0, 25000, by = 5000), 
                     format = "f", digits = 0),
    range = c(0, 20)) +
  theme_void()
  theme(legend.background = element_blank(), legend.position = "right")

```